Preprocess

### 2024-04-15 src/plots.Rmd
Add general age range perf plots

# Bulk calc per model based on age, physician coding stage, and cause of death
calc_by <- function(
        df,
        func,
        name = "Metric",
        by_age = TRUE,
        by_stage = TRUE,
        by_sex = TRUE,
        by_cod = TRUE,
        by_age_range = TRUE,
        ...
    ) {
    out <- list()

@@

    # Calc metric by age range
    if (by_age_range) {

@@

@@

```{r}
df <- raw_df %>% mutate(
    "age_range" = str_to_title(str_replace(
        age_range,
        "\\s*\\(.*\\)",
        ""
    )),
    "age_range" = if_else(
        !str_detect(age_range, "Year|Years|Day|Days|month|Months|week|Weeks"),
        paste0(age_range, " Years"),
        age_range
    )
)
df
```

@@

#### Age Range Performance

@@

@@

>>>>

# Bulk calc per model based on age, physician coding stage, and cause of death
calc_by <- function(
        df,
        func,
        name = "Metric",
        by_age = TRUE,
        by_stage = TRUE,
        by_sex = TRUE,
        by_cod = TRUE,
        by_age_range = TRUE,
        by_age_range2 = TRUE,
        ...
    ) {
    out <- list()

@@

    # Calc metric by detailed age range
    if (by_age_range) {

@@

    # Calc metric by general age range
    if (by_age_range2) {

        # Get unique age ranges
        age_ranges <- df %>%
            pull(age_range2) %>%
            unique

                # Filter for age range
                df_filter <- df %>% filter(
                    age_range2 == ar & is_agreed == TRUE & age == a
                )
                metric_name <- sprintf(
                    "%s %s Age2 Agree %s", name, str_to_title(a), ar
                )

@@

```{r}
df <- raw_df %>% mutate(
    "age_range" = str_to_title(str_replace(
        age_range,
        "\\s*\\(.*\\)",
        ""
    )),
    "age_range" = if_else(
        !str_detect(age_range, "Year|Years|Day|Days|month|Months|week|Weeks"),
        paste0(age_range, " Years"),
        age_range
    ),
    "age_range2" = str_to_title(str_replace(
        age_range2,
        "\\s*\\(.*\\)",
        ""
    )),
    "age_range2" = if_else(
        !str_detect(age_range2, "Year|Years|Day|Days|month|Months|week|Weeks"),
        paste0(age_range2, " Years"),
        age_range2
    )
)
df
```

@@

#### Detailed Age Range Performance

@@

#### General Age Range Performance

```{r}
for (a in c("adult", "neo")) {

    # Get unique age range columns
    arange_col <- out %>% select(starts_with(sprintf(
            "PCCC %s Age2 Agree ",
            aref
        ))) %>%
        names

        # Get cod without prefix
        ar <- gsub(sprintf("PCCC %s Age2 Agree ", aref), "", ar_col)

        # Calc num and perc cases for age range
        narange <- df %>% filter(age_range2 == ar & is_agreed == TRUE & age == a) %>% nrow
        parange <- (narange / arows) * 100
        parange_label <- if (round(parange) >= 1) round(parange) else "<1"

    # Prepare boxplot data
    ar2data <- out %>%
        select(
            Model,
            starts_with(sprintf("PCCC %s Age2 Agree", aref))
        ) %>%
        rename_at(
             vars(-Model),
             ~gsub(sprintf("PCCC %s Age2 Agree ", aref), "", .)
        ) %>%
        filter(!str_detect(Model, "&")) %>% # remove multi models
        pivot_longer( # transform to long format
            cols = -Model,
            names_to = "Age Range",
            values_to = "PCCC"
        ) %>%
        mutate( # rename cod with counts
            "Age Range" = recode(`Age Range`, !!!arange_remap)
        ) %>%
        group_by(`Age Range`) %>%
        mutate( # Make values less than 0 equal to 0
            PCCC = if_else(PCCC <= 0, 0, PCCC)
        ) %>%
        mutate( # Add min, mid, and max model names to label
            "PCCC Min" = if_else(
                PCCC <= min(PCCC) + 0.05,
                paste0(
                    Model,
                    if_else(round(PCCC, 2) != round(min(PCCC), 2), paste0(" (", round(PCCC, 2), ")"), "")
                ),
                NA
            ),
            "PCCC Max" = if_else(
                PCCC >= max(PCCC) - 0.05,
                paste0(
                    Model,
                    if_else(round(PCCC, 2) != round(max(PCCC), 2), paste0(" (", round(PCCC, 2), ")"), "")
                ),
                NA
            ),
            "PCCC Mid" = if_else(
                is.na(`PCCC Min`) & is.na(`PCCC Max`),
                paste0(Model, "\n(", round(PCCC, 2), ")"),
                NA
            ),
            "PCCC Min" = if_else( # Combine into one row if close PCCC
                `PCCC` == min(PCCC),
                if_else(
                    sum(!is.na(`PCCC Min`)) > 1,
                    paste0(na.omit(`PCCC Min`), collapse = "\n"),
                    paste0(
                        str_replace(na.omit(`PCCC Min`), "\\s*\\(.*\\)", ""),
                        collapse = "\n"
                    )
                ),
                NA
            ),
            "PCCC Max" = if_else(
                `PCCC` == max(PCCC),
                if_else(
                    sum(!is.na(`PCCC Max`)) > 1,
                    paste0(na.omit(`PCCC Max`), collapse = "\n"),
                    paste0(str_replace(na.omit(`PCCC Max`), "\\s*\\(.*\\)", ""),
                           collapse = "\n")
                ),
                NA
            )
        )

    # Create cod order based on max pccc
    ar2order <- ar2data %>%
        group_by(`Age Range`) %>%
        summarise("PCCC Max Value" = max(PCCC, na.rm = TRUE)) %>%
        select(`Age Range`, `PCCC Max Value`) %>%
        arrange(desc(`PCCC Max Value`)) %>%
        pull(`Age Range`)
    ardata$`Age Range` <- factor(ardata$`Age Range`, levels = rev(ar2order))

    # Plot boxplot
    ar2plot <- ar2data %>%
        ggplot(aes(x = `Age Range`, y = PCCC)) +
        geom_boxplot(
            linewidth = 0.5,
            width = 0.25
        ) +
        geom_text(
            aes(label = `PCCC Min`),
            color = "#4d4d4d",
            size = 2.5,
            hjust = 1,
            position = position_nudge(y = -0.08)
        ) +
        geom_text(
            aes(label = `PCCC Max`),
            color = "#4d4d4d",
            size = 2.5,
            hjust = 0,
            position = position_nudge(y = 0.08)
        ) +
        stat_summary( # min pccc txt on boxplot
            geom = "text",
            fun = min,
            aes(label = sprintf("%1.2f", after_stat(y))),
            position = position_nudge(y = -0.04),
            size = 3
        ) +
        stat_summary( # max pccc txt on boxplot
            geom = "text",
            fun = max,
            aes(label = sprintf("%1.2f", after_stat(y))),
            position = position_nudge(y = 0.04),
            size = 3
        ) +
        labs(
            x = atitle,
            y = "PCCC (0=Low, 1=High)"
        ) +
        ylim(0.1, 1) +
        coord_flip() +
        theme_minimal() +
        theme(
            panel.grid = element_blank(),
            axis.line = element_blank(),
            axis.line.y = element_blank(),
            axis.title.x = element_text(margin = margin(t = 10)),
            axis.title.y = element_text(margin = margin(r = 10)),
            axis.line.x = element_line(
                color = "black",
                arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
            )
        )

    # Save the plot
    print(ar2plot)
    ggsave(
        sprintf("../manuscript/fig-perf-agerange2-%s.pdf", a),
        plot = ar2plot,
        dpi = 300,
        width = if (a == "adult") 6 else if (a == "neo") 6 else NA,
        height = if (a == "adult") 8 else if (a == "neo") 2.5 else NA
    )
}
```

@@

Remove causes of death with less than 10 samples.

### 2024-04-15 src/preprocess.Rmd
Add general age range perf plots

@@

            "../tmp/healsl_rd1_adult_v1.csv",

@@

            "../tmp/healsl_rd2_adult_v1.csv",

@@

            "../tmp/healsl_rd1_child_v1.csv",

@@

            "../tmp/healsl_rd2_child_v1.csv",

@@

            "../tmp/healsl_rd1_neo_v1.csv",

@@

            "../tmp/healsl_rd2_neo_v1.csv",

@@

@@

```{r}
df$physician <- df$physician %>%
    mutate(
        physician_icd10 = coalesce(
            adj_icd,
            p1_recon_icd,
            p1_icd
        ),
        is_agreed = if_else(
            is.na(p1_recon_icd) & is.na(p2_recon_icd) & is.na(adj_icd),
            TRUE,
            FALSE
        ),
        is_recon = if_else(
            !is.na(p1_recon_icd) | !is.na(p2_recon_icd),
            TRUE,
            FALSE
        ),
        is_adj = if_else(
            !is.na(adj_icd),
            TRUE,
            FALSE
        )
    ) %>%
    select(
        rowid,
        round,
        age,
        age_range,
        sex,
        p1_icd,
        p2_icd,
        p1_recon_icd,
        p2_recon_icd,
        adj_icd,
        is_agreed,
        is_recon,
        is_adj,
        physician_icd10
    )
head(df$physician)
```

@@

>>>>

# Physician age cols
phy_cols_age <- c(
    "rowid",
    "age_value_cod",
    "age_unit_cod"
)

@@

            "../tmp/healsl_rd1_adult_v2.csv",

@@

            "../tmp/healsl_rd2_adult_v2.csv",

@@

            "../tmp/healsl_rd1_child_v2.csv",

@@

            "../tmp/healsl_rd2_child_v2.csv",

@@

            "../tmp/healsl_rd1_neo_v2.csv",

@@

            "../tmp/healsl_rd2_neo_v2.csv",

@@

    ) %>% left_join(
        bind_rows(
                "../tmp/healsl_rd1_adult_age_v2.csv",
                col_select = all_of(phy_cols_age)
                "../tmp/healsl_rd1_child_age_v2.csv",
                col_select = all_of(phy_cols_age)
                "../tmp/healsl_rd1_neo_age_v2.csv",
                col_select = all_of(phy_cols_age)
                "../tmp/healsl_rd2_adult_age_v2.csv",
                col_select = all_of(phy_cols_age)
                "../tmp/healsl_rd2_child_age_v2.csv",
                col_select = all_of(phy_cols_age)
                "../tmp/healsl_rd2_neo_age_v2.csv",
                col_select = all_of(phy_cols_age)
        ) %>% rename(
            age_value = age_value_cod,
            age_unit = age_unit_cod
        by = "rowid"

@@

```{r}
df$physician <- df$physician %>%
    mutate(
        physician_icd10 = coalesce(
            adj_icd,
            p1_recon_icd,
            p1_icd
        ),
        is_agreed = if_else(
            is.na(p1_recon_icd) & is.na(p2_recon_icd) & is.na(adj_icd),
            TRUE,
            FALSE
        ),
        is_recon = if_else(
            !is.na(p1_recon_icd) | !is.na(p2_recon_icd),
            TRUE,
            FALSE
        ),
        is_adj = if_else(
            !is.na(adj_icd),
            TRUE,
            FALSE
        )
    ) %>%
    select(
        rowid,
        round,
        age,
        age_range,
        age_value,
        age_unit,
        sex,
        p1_icd,
        p2_icd,
        p1_recon_icd,
        p2_recon_icd,
        adj_icd,
        is_agreed,
        is_recon,
        is_adj,
        physician_icd10
    )
head(df$physician)
```

@@

### Generalize Age Range

Manually generalize age ranges for each age group:

```{r}
out <- out %>% mutate(
    age_range2 = case_when(
        age_unit == "Days" &
        age_value >= 0 &
        age_value <= 1 ~
            "0-1 Days",
        age_unit == "Days" &
        age_value >= 2 &
        age_value <= 6 ~
            "2-6 Days",
        age_unit == "Days" &
        age_value >= 7 &
        age_value <= 24 ~
            "7-24 Days",
        age_unit == "Years" &
        age_value >= 12 &
        age_value <= 24 ~
            "12-24",
        age_unit == "Years" &
        age_value >= 25 &
        age_value <= 39 ~
            "25-39",
        age_unit == "Years" &
        age_value >= 40 &
        age_value <= 54 ~
            "40-54",
        age_unit == "Years" &
        age_value >= 55 &
        age_value <= 69 ~
            "55-69",
        .default = NA
    ),
    .after = age_range
) %>%
    select(
        -age_value,
        -age_unit
    )
